---
title: "PTR Testes Figures"
author: "Megan Elcheikhali, Advised by Dr. Alex Franks"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(GO.db)
library(tidyverse)
library(patchwork)
library(colorspace)
library(viridis)
library(ggridges)
source("scripts/plot_functions.R")
source("scripts/go_functions.R")
clusters = c("EC", "PTM", "LC", "SPG", "SPC", "St")

mrna_colors = "#B2182B"
protein_colors = "#2166AC"
multi_colors = "lightgray"
mrna_ppc_colors = "#FFBEC1"
protein_ppc_colors = "#C3DBFD"
  
title_size = 80
tag_size = 50
text_size = 80
text_size_sub = 55
legend_size = 3.5
```

## Model Overview

$M_{gc}^d \overset{\triangle}{=}$ Sum of transcript counts for gene $g$ across cells associated with cell type $c$ in data set $d$. 

$P_{g_kc}^d \overset{\triangle}{=}$ Average $\text{log}_2$ intensity for peptide $k$ associated with protein $g$ across cells associated with cell type $c$ in data set $d$. 

$M_{gc}^d \sim$ Negative Binomial$(2^{\mu_{gc}} \gamma_{gd} a_{cd} ,\phi_d)$

$P_{g_kc}^{d} \sim$ Normal$(s_g(\mu_{gc} + R_{gc}) + \kappa_{g, d} + b_{cd}, \sigma^2_{gd})$

Where, 

$\mu_{gc} \overset{\triangle}{=}$ Underlying mRNA intensity in gene $g$ and cell type $c$.

$R_{gc} \overset{\triangle}{=}$ rPTR in gene $g$ and cell type $c$.

$\gamma_{gd} \overset{\triangle}{=}$ mRNA technical effect for gene $g$, data set $d$.

$\kappa_{gd} \overset{\triangle}{=}$ Protein technical effect for gene $g$, data set $d$.

$a_{cd} \overset{\triangle}{=}$ mRNA technical effect for cell type $c$, data set $d$.

$b_{cd} \overset{\triangle}{=}$ Protein technical effect for cell type $c$, data set $d$.

$\phi_d \overset{\triangle}{=}$ mRNA overdispersion parameter for data set $d$.

$\sigma_{gd}^2 \overset{\triangle}{=}$ Protein variance parameter term for data set $d$ and protein $g$.

```{r fig2, echo = FALSE, warning = FALSE, message = FALSE, fig.width = 35, fig.height = 15, cache = F}
# FIGURE 2: Empirical Across Cluster Correlations
# load previously computed correlations
load("processed_data/observed_z.RData")
mrna_pop1_z = mrna_pop1_z %>% mutate(pop_mrna = "Pop1") %>% filter(ct %in% clusters & is.finite(mrna_zscore)) 
mrna_pop2_z = mrna_pop2_z %>% mutate(pop_mrna = "Pop2") %>% filter(ct %in% clusters & is.finite(mrna_zscore)) 
mrna_prot_rec = AnnotationDbi::select(org.Hs.eg.db, 
                                      keys = intersect(mrna_pop1_z$SYMBOL, mrna_pop2_z$SYMBOL), 
                                      keytype = "SYMBOL", columns = "UNIPROT")

protein_pop1_z = protein_pop1_z %>% mutate(pop_protein = "Pop1") %>% filter(ct %in% clusters & is.finite(protein_zscore)) 
protein_pop2_z = protein_pop2_z %>% mutate(pop_protein = "Pop2") %>% filter(ct %in% clusters & is.finite(protein_zscore)) 
protein_pop4_z = protein_pop4_z %>% mutate(pop_protein = "Pop4") %>% filter(ct %in% clusters & is.finite(protein_zscore))

load("processed_data/p1_p2_ac_cor.RData")
cor_df_prot = cor_df %>% filter(n_ct >= 3)

load("processed_data/p1_p4_ac_cor.RData")
cor_df = cor_df %>% filter(n_ct >=3)
cor_df_prot = cor_df_prot %>% rbind(cor_df)

load("processed_data/p2_p4_ac_cor.RData")
cor_df = cor_df %>% filter(n_ct >= 3)
cor_df_prot = cor_df_prot %>% rbind(cor_df)

load("processed_data/protein_reliability_ct.RData")
group_ref = protein_reliability %>%
  stats::na.omit() %>%
  filter(pop_protein != "Pop4") %>%
  dplyr::group_by(UNIPROT) %>%
  dplyr::summarise(rel_av = mean(reliability, na.rm = T)) %>%
  ungroup() %>%
  dplyr::mutate(protein_threshold = quantile(rel_av, 0.90, na.rm = T),
                thresh = rel_av >= protein_threshold) %>%
  dplyr::group_by(UNIPROT) %>%
  dplyr::summarise(var_group = ifelse(thresh, "Group A", "Group B"))
group_ref_sub = data.frame(UNIPROT = setdiff(unique(cor_df_prot$UNIPROT), unique(group_ref$UNIPROT)), var_group = "Group B")
group_ref = group_ref %>% rbind(group_ref_sub)

load("processed_data/m1_m2_ac_cor.RData") 
mrna_cor_df = cor_df %>% filter(n_ct >= 3)

load("processed_data/m1_p1_ac_cor.RData")
cor_df_multi = cor_df %>% filter(n_ct >= 3)

load("processed_data/m2_p1_ac_cor.RData")
cor_df = cor_df %>% filter(n_ct >= 3)
cor_df_multi = cor_df_multi %>% rbind(cor_df)

load("processed_data/m1_p2_ac_cor.RData")
cor_df = cor_df %>% filter(n_ct >= 3)
cor_df_multi = cor_df_multi %>% rbind(cor_df)

load("processed_data/m2_p2_ac_cor.RData")
cor_df = cor_df %>% filter(n_ct >= 3)
cor_df_multi = cor_df_multi %>% rbind(cor_df)

load("processed_data/m1_p4_ac_cor.RData")
cor_df = cor_df  %>% filter(n_ct >= 3)
cor_df_multi = cor_df_multi %>% rbind(cor_df)

load("processed_data/m2_p4_ac_cor.RData")
cor_df = cor_df %>% filter(n_ct >= 3)
cor_df_multi = cor_df_multi %>% rbind(cor_df)

p = plot_across_clusters_corr_joined(mrna_cor_df = mrna_cor_df, prot_cor_df = cor_df_prot, multi_cor_df = cor_df_multi, 
                                     color_select = c(mrna_colors, protein_colors, "lightgray"), group_ref = group_ref) +
  xlab("") +
  ylim(-1, 1)
```

```{r fig2b, echo = FALSE, warning = FALSE, message = FALSE, fig.width = 35, fig.height = 30, cache = F, eval = T}
load("processed_data/mrna_cor_cors_ab.RData")
load("processed_data/protein_cor_cors_ab.RData")
load("processed_data/multi_modal_cor_cors_ab.RData")
symbols = c(as.character(unique(mrna_cor_cors_a$gene1)), as.character(unique(mrna_cor_cors_b$gene1)))
mrna_ref = AnnotationDbi::select(org.Hs.eg.db, keys = symbols, 
                                 keytype = "SYMBOL", columns = c("UNIPROT"))
mrna_cor_cors = mrna_cor_cors_a %>% rbind(mrna_cor_cors_b) %>% 
  dplyr::mutate(SYMBOL = gene1) %>% merge(mrna_ref) %>% dplyr::select(-c(gene1, SYMBOL))
protein_cor_cors = protein_cor_cors_a %>% rbind(protein_cor_cors_b) %>% 
  mutate(UNIPROT = gene1) %>% dplyr::select(-gene1)
multi_modal_cor_cors = multi_modal_cor_cors_a %>% rbind(multi_modal_cor_cors_b) %>% 
  mutate(UNIPROT = gene1) %>% dplyr::select(-gene1)

q = plot_across_clusters_corr_joined(mrna_cor_df = mrna_cor_cors, prot_cor_df = protein_cor_cors,
                                     multi_cor_df = multi_modal_cor_cors, 
                                     color_select = c(mrna_colors, protein_colors, "lightgray"), 
                                     group_ref = group_ref) +
  xlab("") +
  ylab("Correlation of Correlation Vectors") +
  ylim(-1, 1)

p / q +
  plot_annotation(title = "Figure 1: Correlation of mRNA, Protein Observed Data",
  theme = theme(plot.title = element_text(size = title_size - 2), plot.tag = element_text(size = tag_size))) &
  theme(text = element_text(size = text_size - 10))
```

\newpage

```{r high, echo = FALSE, warning = FALSE, message = FALSE, fig.width = 15, fig.height = 8, cache = F}
load("model_output_hs/gene_res_centered.RData")
genes = c("P22087", "P17174")
title_lab = c("High Variance, High Agreement", "High Variance, Low Agreement")
p1 = plot_zscore_means_alone(mrna_z = list(mrna_pop1_z, mrna_pop2_z), 
                             protein_z = list(protein_pop1_z, protein_pop2_z, protein_pop3_z, protein_pop4_z), 
                             gene_res = gene_res,
                             uni = genes[1], clusters = clusters, 
                             mrna_pop_label = c("Pop1", "Pop2"), protein_pop_label = c("Pop1", "Pop2", "Pop3", "Pop4"),
                             mrna_color = mrna_colors, protein_color = protein_colors, 
                             title_select = title_lab[1]) + 
  ylim(-2, 2)

p2 = plot_zscore_means_alone(mrna_z = list(mrna_pop1_z, mrna_pop2_z), 
                             protein_z = list(protein_pop1_z, protein_pop2_z, protein_pop3_z, protein_pop4_z), 
                             gene_res = gene_res,
                             uni = genes[2], clusters = clusters, 
                             mrna_pop_label = c("Pop1", "Pop2"), protein_pop_label = c("Pop1", "Pop2", "Pop3", "Pop4"),
                             mrna_color = mrna_colors, protein_color = protein_colors, 
                             title_select = title_lab[2]) + 
  ylim(-2, 2)


p1 + p2 + plot_layout(guides = "collect") +
  plot_annotation(title = "Figure 3: Observed LFCs with High Variation and High, Low Agreement",
                  tag_levels = 'A', theme = theme(plot.title = element_text(size = 30), 
                                                  plot.tag = element_text(size = 30))) &
    theme(text = element_text(size = 35),
          title = element_text(size = 25),
          legend.position = "bottom")
```

\newpage

```{r fig3, echo = FALSE, warning = FALSE, message = FALSE, fig.width = 5, fig.height = 6, cache = F}
# FIGURE 3: Posterior Predictive Model Evaluations
# load previously completed variance quantities
load("model_output_hs/mrna_ppc_var.RData")
m1 = plot_mrna_ppc_variance(ppc_var_plot = mrna_variance_ppc, rescale = F)

load("model_output_hs/protein_ppc_var.RData")
p1 = plot_protein_ppc_variance(ppc_var_plot = protein_variance_ppc) + theme(legend.position = "none")

load("model_output_hs/ppc_cor.RData")
mp = draw_ppc_correlation_density(ppc_cor_obj, mrna_df = "Pop1")

p = m1 / mp + plot_layout(guides = "collect") 
p + plot_annotation(title = "Figure 4: Posterior Predictive and Observed Statistics",
  tag_levels = 'A') & 
  theme(plot.title = element_text(size = 14), plot.tag = element_text(size = 12), 
        legend.position = "bottom",
        text = element_text(size = 12),
        axis.text = element_text(size = 12))
```

\newpage

```{r fig4, echo = FALSE, warning = FALSE, message = FALSE, fig.width = 20, fig.height = 20, cache = F}
# FIGURE 4: Evaluating posterior samples of mu and r
m1 = plot_posterior_means(gene_res = gene_res) + theme(legend.position = "none",
                                                       text = element_text(size = 40),
                                                       axis.text = element_text(size = 40))

load("model_output_hs/sig_intervals.RData")
ct_select = "PTM"
uni_select = "P68104"
p1 = plot_significant_intervals(sig_intervals = significant_intervals, gene_res = gene_res, 
                                ct_select = ct_select, organism = "human", sig_color = "purple") + 
  theme(text = element_text(size = 40),
        axis.text = element_text(size = 40))

load("processed_data/observed_z.RData")
load("model_output_hs/mrna_ppc_z.RData")
load("model_output_hs/protein_ppc_z.RData")
# mrna_ppc_z = mrna_ppc_z %>%
#   dplyr::group_by(pop_mrna, ct) %>%
#   dplyr::mutate(mrna_zscore = mrna_zscore - mean(mrna_zscore, na.rm = T))
# protein_ppc_z = protein_ppc_z %>%
#   dplyr::group_by(pop_protein, ct) %>%
#   dplyr::mutate(protein_zscore = protein_zscore - mean(protein_zscore, na.rm = T))

p2 = plot_zscore_means(mrna_z = list(mrna_pop1_z, mrna_pop2_z),
                       protein_z = list(protein_pop1_z, protein_pop2_z, protein_pop3_z, protein_pop4_z),
                       gene_res = gene_res,
                       mrna_ppc_z = mrna_ppc_z,
                       protein_ppc_z = protein_ppc_z,
                       uni = uni_select, clusters = clusters,
                       mrna_pop_label = c("Pop1", "Pop2"), protein_pop_label = c("Pop1", "Pop2", "Pop3", "Pop4"),
                       mrna_color = mrna_colors, protein_color = protein_colors,
                       mrna_ppc_color = mrna_ppc_colors, protein_ppc_color = protein_ppc_colors)

load("model_output_hs/cor_comparison.RData")
mp = plot_correlation_comparison_med(cor_comparison = cor_comparison, gene_res = gene_res) 

design = "
AAAA#
CCCCD
CCCCD
CCCCD
CCCCD
"
mp2 = (mp[[1]][[1]] + mp[[1]][[2]] + mp[[2]][[2]]) + 
  plot_layout(design = design, width = c(4, 4, 1), height = c(1, 4, 4))

p = m1 / 
    mp2 | #& theme(text = element_text(size = 40), 
                # plot.margin = unit(0, "cm"),
                # panel.margin = unit(0, "cm"),
                # legend.key.size = unit(legend_size, "cm"),
                # legend.position = "bottom") | 
   wrap_elements(full = p2 & theme(plot.title = element_text(hjust = 0.5),
                                   text = element_text(size = 40),
                                   legend.key.size = unit(2, "cm"),
                                   legend.position = "bottom")) +
  plot_layout(guides = "collect")
p + 
  plot_annotation(tag_levels = list(c("A", "", "B", "", "C")),
  title = "Figure 5: Evaluation of Posterior Samples and Gene-Level rPTR",
  theme = theme(plot.title = element_text(size = 50), 
                plot.tag = element_text(size = 40, hjust = 0, vjust = 0),
                plot.tag.position = c(0, 1)))  &
   theme(text = element_text(size = 40),
         legend.key.size = unit(2, "cm"),
         legend.position = "bottom")
```

\newpage

```{r fig5, echo = FALSE, warning = FALSE, message = FALSE, fig.width = 35, fig.height = 40, eval = T, cache = F}
load("model_output_hs/go_test_results_centered.RData")

# GO_heatmap = c("GO:0035686", "GO:0097513", "GO:0051414",
#                "GO:0030020", "GO:0030317", "GO:0045071", 
#                "GO:0071230", "GO:0007286", "GO:0000027", "GO:0098641")

GO_heatmap = c("GO:0035686", "GO:0030317", "GO:0005853", "GO:0030317", "GO:0035686",
               "GO:0090131", "GO:0071230", "GO:0042788", "GO:0048407", "GO:0090131")


GO_ec = c("GO:0008307", "GO:0006614", "GO:0006414", "GO:0031526", "GO:0015935")
GO_ptm = c("GO:0007286", "GO:0030020", "GO:0030317", "GO:0061631", "GO:0071230")
GO_lc = c("GO:0007018", "GO:0007286", "GO:0001568", "GO:0030020", "GO:0002181")
GO_spg = c("GO:0001568", "GO:0030048", "GO:0030240", "GO:0001568", "GO:0003009")
GO_st = c("GO:0002080", "GO:0003777", "GO:0007254", "GO:0048240", "GO:0097228")
#######################
b = plot_go_heatmap(test_res, go_select = GO_heatmap) + 
  theme(text = element_text(size = text_size_sub),
        axis.text.x = element_text(size = text_size_sub, angle = 60, vjust = 1, hjust=1),
        legend.position = "bottom",
        legend.key.size = unit(legend_size, "cm"))

celltype_title = c("Endothelial", "Peritubular Myoid", "Leydig", "Spermatagonia", "Spermatocyte", "Spermatid")

load(paste0("model_output_hs/GO_dat_", clusters[3], ".RData"))
d = plot_ticks2(GO_dat = GO_dat, gene_res = gene_res, GO_list = GO_lc,
                    celltype = clusters[3], celltype_title = celltype_title[3]) 

load(paste0("model_output_hs/GO_dat_", clusters[6], ".RData"))
f = plot_ticks2(GO_dat = GO_dat, gene_res = gene_res, GO_list = GO_st,
                    celltype = clusters[6], celltype_title = celltype_title[6]) 

design <- "
  AAABCCCD
  EEEFGGGH
  EEEFGGGH
  EEEFGGGH
  EEEFGGGH
  EEEFGGGH
  EEEFGGGH
  EEEFGGGH
  EEEFGGGH
  ########
  IIIIIIII
  IIIIIIII
  IIIIIIII
  IIIIIIII
  IIIIIIII
  IIIIIIII
  IIIIIIII
  IIIIIIII
  IIIIIIII
  IIIIIIII
  IIIIIIII
  IIIIIIII
"

# q = wrap_elements(full = b) + 
#     plot_annotation(tag_levels = list(c("A")), 
#                     title = "Figure 4: Evaluation of GO Groups with Significant rPTR",
#                     theme = theme(plot.title = element_text(size = 50))) &
#   theme(plot.tag = element_text(size = 50))
  
p = d[[1]][[1]] + d[[2]][[1]] + f[[1]][[1]] + f[[2]][[1]] + d[[1]][[2]] + d[[2]][[2]] + f[[1]][[2]] + f[[2]][[2]] + wrap_elements(full = b) +
    plot_layout(design = design, guides = "collect") 
p = p + plot_annotation(tag_levels = list(c("", "", "", "", "A", "", "B", "", "C")),
                        title = "Figure 6: Evaluation of GO Groups with Significant rPTR",
                        theme = theme(plot.title = element_text(size = title_size), 
                                       plot.tag = element_text(size = tag_size))) &
    theme(text = element_text(size = text_size_sub + 4),
          axis.title.x = element_text(size = text_size_sub - 8),
          legend.position = "bottom", 
          legend.key.size = unit(legend_size, "cm"))

p
```
